Comparison of several sequential and parallel derivative-free global optimization algorithms

### 2018-12-01 2019/PAVT/pavt_paper.tex
Some minor corrections

The GKLS generator \cite{Gaviano2003} allows obtaining the problems of given dimensionality
with given number of extrema. Besides, GKLS allows adjusting the complexity of the problems by
decreasing or increasing the size of the global minimum attractor. In Ref.
\cite{SergeyevKvasov2006}, the parameters of the generator allowing generating the sets of 100
problems each of two levels of complexity (Simple and Hard) of the dimensionality equal to 2, 3, 4,
and 5 are given. Following the authors of the GKLS generator, we will use the parameters proposed
by them and, this way, add 800 more problems of various dimensionalities and complexity into the
test problem set.

Let us suppose a test problem to be solved if the optimization method executes the scheduled trial
\(y^k\) in a \(\delta\)-nearness of the global minimum \(y^*\) i. e. $\left\|y^k-y^*\right\|\leq \delta
= 0.01\left\|b-a\right\|$ where \(a\) and \(b\) are the left and the right boundaries of the hypercube
from (\ref{eq:task}). If this relation is not fulfilled before the expiration of the limit of the number of
trials, the problem was considered to be unsolved. The maximum limit of the number of trials was set
for each problem class according to the dimensionality and complexity (see Table \ref{tab:limits}).

@@

\section{Results of Numerical Experiments}
\subsection{Results of Sequential Algorithms}

The results of various algorithms on different problem classes depend on the adjustments of
algorithms directly. In most cases, the authors of software implementations are oriented onto the
problems of medium difficulty. In order to obtain a satisfactory result when solving the essentially
multiextremal problems, a correction of some parameters is required. When conducting the
comparison, the following parameters for the methods were employed:
\begin{itemize}
  \item in the AGS\(l\) method, the parameter of alternation of the global rules of computing the
characteristic and the local one was set to be equal to 5:1;
  \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
  \item in the SDA method, the parameter \(visit=2.72\).
\end{itemize}

@@

\subsection{Results of Parallel Algorithms}

@@

In the present paper, several sequential and parallel global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
making the following conclusions:
\begin{itemize}
  \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
speed and reliability at the level of \(DIRECT\) and exceeds many other algorithms, the open-access
implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
  \item the parallel version of the Globalizer system demonstrates good indicators of speedup in time
when running on several nodes, on each of which a single objective function value per iterations is
computed. When solving the problems with fast computable objective functions and using multiple
threads on the nodes, the indicators of speedup for the Globalizer system degrade with increasing
number of nodes;
  \item the MIDACO system is the most suitable for simple global optimization problems with the
fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
speedup with increasing number of nodes and threads executed on these ones in parallel.
\end{itemize}

\section*{Acknowledgements}
The study was supported by the Russian Science Foundation, project No 16-11-
10150.

>>>>

The GKLS generator \cite{Gaviano2003} allows obtaining the problems of given dimensionality
with given number of extrema. Moreover, GKLS allows adjusting the complexity of the problems by
decreasing or increasing the size of the global minimum attractor. In 
\cite{SergeyevKvasov2006} the parameters of the generator allowing generating the sets of 100
problems each of two levels of complexity (Simple and Hard) of the dimensionality equal to 2, 3, 4,
and 5 are given. Following the authors of the GKLS generator, we will use the parameters proposed
by them and, this way, add 800 more problems of various dimensionalities and complexity into the
test problem set.

Let us suppose a test problem to be solved if the optimization method executes the scheduled trial
\(y^k\) in a \(\delta\)-vicinity of the global minimum \(y^*\), i.e. $\left\|y^k-y^*\right\|\leq \delta
= 0.01\left\|b-a\right\|$, where \(a\) and \(b\) are the left and the right boundaries of the hypercube
from (\ref{eq:task}). If this relation is not fulfilled before the expiration of the limit of the number of
trials, the problem was considered to be unsolved. The limit of the number of trials was set
for each problem class according to the dimensionality and complexity (see Table \ref{tab:limits}).

@@

\section{Results of Numerical Experiments}
\subsection{Sequential Algorithms}

The results of various algorithms on different problem classes depend on the adjustments of
algorithms directly. In most cases, the authors of software implementations are oriented onto the
problems of medium difficulty. In order to obtain a satisfactory result when solving the essentially
multiextremal problems, a correction of some parameters is required. When conducting the
comparison, the following parameters for the methods were employed:
\begin{itemize}
  \item in the AGS\(l\) method, the parameter of alternation the
global and local iterations was set to be equal to 5:1;
  \item in the DIRECT and DIRECT\(l\) methods, the parameter \(\epsilon=10^{-4}\);
  \item in the SDA method, the parameter \(visit=2.72\).
\end{itemize}

@@

\subsection{Parallel Algorithms}

@@

In the present paper, several sequential and parallel global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
making the following conclusions:
\begin{itemize}
  \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-access
implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
  \item the parallel version of the Globalizer system demonstrates good speedup
when running on several nodes, on each of which a single objective function value per iterations is
computed. When solving the problems with fast computable objective functions and using multiple
threads on the nodes, the speedup for the Globalizer system degrade with increasing
number of nodes;
  \item the MIDACO system is the most suitable for simple global optimization problems with the
fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
speedup with increasing number of nodes and threads executed on these ones in parallel.
\end{itemize}

\section*{Acknowledgements}
The study was supported by the Russian Science Foundation, project No 16-11-10150.
